2024 US Presidential Election Forecast Model

### 2024-11-04 paper/paper.qmd
update paper.qmd

---
title: "2024 US Presidential Election Forecast Model"
subtitle: "My subtitle if needed"
author: 
  - Viet Nguyen
  - Yihang Xu
  - Doran Wang
thanks: "Code and data are available at: [https://github.com/vietng04/2024-US-Presidential-Election-Forecast](https://github.com/vietng04/2024-US-Presidential-Election-Forecast)."
date: today
date-format: long 
abstract: "This paper uses aggregate polling data and 'STH' modeling methods to predict the 2024 U.S. presidential election outcome. The Model use the factors related to the polls on Donald Trump’s support rate. The study also includes a deep-dive analysis of methodology on selected pollster, the New York Times (NYT), and then an idealized methodology and survey for predicting elections on a limited budget."
format: pdf
toc: true
toc-title: "Table of Contents"
number-sections: true
bibliography: references.bib
---

@@

# Data {#sec-data}
The data used in this paper came from the FiveThirtyEight [@fivethirtyeight2024]. Data were cleaned and analyzed using the open source statistical programming language R [@citeR]. Libraries `tidyverse` [@citeTidyverse], `janitor` [@Citejan], `knitr` [@citeknitr], and `dplyr` [@citedplyr] were used for simulating, cleaning and testing. Graphics were made using `ggplot2` [@citeGG].

@@

The @tbl-cleaned_data shows the 5 of all colunms.

@@

```{r}
#| label: fig-distribution_Answers
#| fig-cap: Distribution of Answers
#| echo: false
#| warning: false
#| message: false

ggplot(cleaned_data %>% filter(answer %in% c("Trump", "Harris")), aes(x = answer)) +
  geom_bar(fill = "pink", color = "pink") +
  labs(x = "Answer", y = "Count") +
  theme_minimal()
```

```{r}
#| label: fig-histogram_pct
#| fig-cap: Percentage of Support for Each Candidate
#| echo: false
#| warning: false
#| message: false

ggplot(cleaned_data %>% filter(answer %in% c("Trump", "Harris")), aes(x = answer, y = pct, fill = party)) +
  geom_bar(stat = "identity", color = "lightblue") +
  labs( x = "Candidate", y = "Percentage") +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) + # Converts y-axis to percentage format
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotates x-axis labels
```

```{r}
#| label: transparency score for each candidate
#| fig-cap: Percentage of Support for Each Candidate
#| echo: false
#| warning: false
#| message: false
ggplot(cleaned_data, aes(x = answer, y = transparency_score, fill = answer)) +
  geom_boxplot() +
  labs(title = "Transparency Score by Candidate", x = "Candidate", y = "Transparency Score") +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}
#| label: fig-pct_overtime
#| fig-cap: Mean Percentage of Support for Trump and Harris over time
#| echo: false
#| warning: false
#| message: false
#| 
trump_data <- cleaned_data %>%
  filter(answer == "Trump") %>%
  mutate(end_date = mdy(end_date)) %>%  # Convert end_date to Date format
  group_by(end_date) %>%
  summarize(mean_pct = mean(pct, na.rm = TRUE))

ggplot(trump_data, aes(x = end_date, y = mean_pct)) +
  geom_smooth(method = "loess", color = "red", fill = "#FF7F7F", linetype = "solid", size = 1.3, se = TRUE, ) +
  annotate("text", 
           x = max(trump_data$end_date), 
           y = trump_data$mean_pct[which.max(trump_data$end_date)], 
           label = paste0("Trump ", round(trump_data$mean_pct[which.max(trump_data$end_date)], 1), "%"),
           color = "red", 
           fontface = "bold",
           size = 2) +
  labs(title = NULL, x = "Date", y = "Percentage") +
  theme_minimal() +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

## Trump
The model estimates the influence of several predictors on the dependent variable. The intercept is estimated at 48.56, indicating the baseline level of the dependent variable when all predictors are held at zero.

@@

## Harris
The model investigates the impact of several predictors on the dependent variable, with the following key findings:

>>>>

---
title: "2024 US Presidential Election Forecast Model"
author: 
  - Viet Nguyen
  - Yihang Xu
  - Doran Wang
thanks: "Code and data are available at: [https://github.com/vietng04/2024-US-Presidential-Election-Forecast](https://github.com/vietng04/2024-US-Presidential-Election-Forecast)."
date: today
date-format: long 
abstract: "This paper uses aggregate polling data and 'STH' modeling methods to predict the 2024 U.S. presidential election outcome. The Model use the factors related to the polls on Donald Trump’s support rate. The study also includes a deep-dive analysis of methodology on selected pollster, the New York Times (NYT), and then an idealized methodology and survey for predicting elections on a limited budget."
format: pdf
toc: true
toc-title: "Table of Contents"
number-sections: true
bibliography: references.bib
---

@@

# Data {#sec-data}
The data used in this paper came from the FiveThirtyEight [@fivethirtyeight2024]. Data were cleaned and analyzed using the open source statistical programming language R [@citeR]. Libraries `tidyverse` [@citeTidyverse], `arrow` [@citearrow] `janitor` [@Citejan], `knitr` [@citeknitr], and `dplyr` [@citedplyr] were used for simulating, cleaning and testing. Graphics were made using `ggplot2` [@citeGG].

@@

@@

```{r}
#| label: fig-distribution_Answers
#| fig-cap: Distribution of Answers
#| fig-width: 2
#| fig-height: 2
#| echo: false
#| warning: false
#| message: false

ggplot(cleaned_data %>% filter(answer %in% c("Trump", "Harris")), aes(x = answer)) +
  geom_bar(fill = "pink", color = "pink") +
  labs(x = "Answer", y = "Count") +
  theme_minimal()
```
```{r}
#| label: fig-pollster_party
#| fig-cap: Count of Polls by Pollster and Party
#| fig-width: 2
#| fig-height: 2
#| echo: false
#| warning: false
#| message: false
filter_data <- cleaned_data %>%
  filter(answer %in% c("Trump", "Harris"))

ggplot(filter_data, aes(x = answer, y = pct)) +
  geom_point(color = "steelblue", alpha = 0.6) +  # Points with some transparency
  labs(title = "Relationship between Answer and Percentage", 
       x = "Answer", 
       y = "Percentage (pct)") +  # Labels for axes and title
  theme_minimal()
```

## Trump
```{r}
#| label: fig-pct_overtime
#| fig-cap: Mean Percentage of Support for Trump over time
#| fig-width: 8
#| fig-height: 3
#| echo: false
#| warning: false
#| message: false

trump_data <- cleaned_data %>%
  filter(answer == "Trump") %>%
  mutate(end_date = mdy(end_date)) %>%  # Convert end_date to Date format
  group_by(end_date) %>%
  summarize(mean_pct = mean(pct, na.rm = TRUE))

ggplot(trump_data, aes(x = end_date, y = mean_pct)) +
  geom_smooth(method = "loess", color = "red", fill = "#FF7F7F", linetype = "solid", size = 1.3, se = TRUE, ) +
  annotate("text", 
           x = max(trump_data$end_date), 
           y = trump_data$mean_pct[which.max(trump_data$end_date)], 
           label = paste0("Trump ", round(trump_data$mean_pct[which.max(trump_data$end_date)], 1), "%"),
           color = "red", 
           fontface = "bold",
           size = 2) +
  labs(title = NULL, x = "Date", y = "Percentage") +
  theme_minimal() +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```
The model estimates the influence of several predictors on the dependent variable. The intercept is estimated at 48.56, indicating the baseline level of the dependent variable when all predictors are held at zero.

@@

## Harris
```{r}
#| label: fig-pct_overtime2
#| fig-cap: Mean Percentage of Support for Harris over time
#| fig-width: 8
#| fig-height: 3
#| echo: false
#| warning: false
#| message: false
harris_data <- cleaned_data %>%
  filter(answer == "Harris") %>%
  mutate(end_date = mdy(end_date)) %>%  # Convert end_date to Date format
  group_by(end_date) %>%
  summarize(mean_pct = mean(pct, na.rm = TRUE))

ggplot(harris_data, aes(x = end_date, y = mean_pct)) +
  geom_smooth(method = "loess", color = "blue", fill = "blue", linetype = "solid", size = 1.3, se = TRUE, ) +
  annotate("text", 
           x = max(harris_data$end_date), 
           y = harris_data$mean_pct[which.max(harris_data$end_date)], 
           label = paste0("Harris ", round(harris_data$mean_pct[which.max(harris_data$end_date)], 1), "%"),
           color = "blue", 
           fontface = "bold",
           size = 2) +
  labs(title = NULL, x = "Date", y = "Percentage") +
  theme_minimal() +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```
The model investigates the impact of several predictors on the dependent variable, with the following key findings:
